### 2017-04-05 icwe-2017-review-response.md
Add reviews

Dear Sarven,

We are very pleased to inform you that your submission entitled

Decentralised Authoring, Annotations and Notifications for a Read-Write-Web with dokieli

has been ACCEPTED as an application paper at ICWE 2017.

** CAMERA READY INSTRUCTION **

Please prepare the camera-ready version of your paper (maximum length is 12 pages) by **April 7th** by addressing the reviewer comments below and follow STRICTLY the Springer LNCS guidelines for authors at http://www.springer.de/comp/lncs/authors.html. This is a HARD deadline, no extensions will be granted. Springer may reject your paper if the manuscript is not properly formatted. The camera-ready version of your paper (sources in a zip + separate pdf) and the copyright form (in pdf, signed and scanned) should be uploaded on EasyChair.

A second email will follow with detailed instructions.

In case any further questions arise during the preparation of your camera-ready version, please do not hesitate to contact the program chairs AND the proceedings Chair (Florian Daniel, florian.daniel at polimi.it).

** CONFERENCE AUTHOR REGISTRATION **

Please do not forget that at least one author per accepted submission needs to register as a CONFERENCE AUTHOR attendee in order for the paper to appear in the proceedings. Information about the registration procedure is available on the conference Web site. If you have multiple papers accepted, each paper must have a separate CONFERENCE AUTHOR registration.

Thank you for choosing ICWE 2017; we look forward to seeing you in Rome!

Kind regards, 
ICWE 2017 Application Track Chair

----------------------- REVIEW 1 ---------------------
PAPER: 152
TITLE: Decentralised Authoring, Annotations and Notifications for a Read-Write-Web with dokieli
AUTHORS: Sarven Capadisli, Amy Guy, Ruben Verborgh, Christoph Lange, Sören Auer and Tim Berners-Lee

----------- Overall evaluation -----------
This paper presents dokieli, a decentralized web-based authoring service. 
The paper is very well-written and a pleasure to read. The service presented here is nice and useful, it defines a new way of publishing and interacting with content on the Web. There are many on-line resources that can be triggered to understand more about this service; indeed, the reader can access the github repository to gather more information and also the dokieli Website is a useful resource. Nevertheless, the paper itself does not provide many technical details and remains quite at a high level.

The contribution is well-suited for the conference and it should generate interesting discussions at the conference.

----------------------- REVIEW 2 ---------------------
PAPER: 152
TITLE: Decentralised Authoring, Annotations and Notifications for a Read-Write-Web with dokieli
AUTHORS: Sarven Capadisli, Amy Guy, Ruben Verborgh, Christoph Lange, Sören Auer and Tim Berners-Lee

----------- Overall evaluation -----------
The authors present dokieli, a  decentralised, browser-based authoring and annotation platform to create “living” documents that are interoperable and independent of dokieli. Dokieli follows standards and best practices, mainly HTML+RDFa for, Linked Data Platform for personal data storage, and Linked Data Notifications for updates. The presented platform has real potential in the context of academic/journalistic writings. The paper presents good arguments around the use of such a decentralised system to offer authors ownership and sovereignty over their data. The implementation is also sound. However, the paper lacks rigour from the perspective of comparing work against existing approaches. The authors do compare and contrast their system with other offerings; however, the research work and techniques used by other is not critically analysed (for example, there are almost no references in the main text. However, there is an excellent list of references at the end, was th!
 is intentional?).

----------------------- REVIEW 3 ---------------------
PAPER: 152
TITLE: Decentralised Authoring, Annotations and Notifications for a Read-Write-Web with dokieli
AUTHORS: Sarven Capadisli, Amy Guy, Ruben Verborgh, Christoph Lange, Sören Auer and Tim Berners-Lee

----------- Overall evaluation -----------
The authors illustrate a decentralized framework for web-based authoring and annotation service.

The paper is really well-written, organized and motivated.

The main weakness is the nature of the proposal more fitting a demonstration context rather than an application/industrial scenario.

However the contribution will leverage a fruitful discussion during the conference.